Unbiased pangenome graphs

### 2022-02-14 manuscript/main.tex
stuff

\abstract{
\textbf{Motivation:}
Pangenome variation graphs model the mutual alignment of collections of DNA sequences.
A set of pairwise alignments implies a variation graph, but there are no scalable methods to generate such a graph from these alignments.
Existing related approaches depend on a single reference, a specific ordering of genomes, or a \textit{de Bruijn} model based on a fixed $k$-mer length. %, and may not describe the base-level alignment of genomes into the pangenome.
A scalable, self-contained method to build pangenome graphs without such limitations would be a key step in pangenome construction and manipulation pipelines. \\
\textbf{Results:}
We design the \textit{seqwish} algorithm, which builds a variation graph from a set of sequences and alignments between them.
We first transform the alignment set into an implicit interval tree.
To build up the variation graph, we query this tree-based representation of the alignments to reduce transitive matches into single DNA segments in a sequence graph.
By recording the mapping from input sequence to output graph, we can trace the original paths through this graph, yielding a pangenome variation graph.
We present an implementation that operates in external memory, using disk-backed data structures and a lock-free parallel algorithm to drive the core graph induction step.
We demonstrate that our method scales to very large graph induction problems by applying it to build pangenome graphs for several species. \\
\textbf{Availability:}
\textit{seqwish} is published as free software under the MIT open source license.
Source code and documentation are available at \url{https://github.com/ekg/seqwish}.
\textit{seqwish} can be installed via Bioconda \url{https://bioconda.github.io/recipes/seqwish/README.html} or GNU Guix \url{https://github.com/ekg/guix-genomics/blob/master/seqwish.scm}. \\
\textbf{Contact:} \href{egarris5@uthsc.edu}{egarris5@uthsc.edu} \\
%\textbf{Supplementary information:} Supplementary data are available at \textit{Bioinformatics} online.
}

>>>>

\abstract{
\textbf{Motivation:}
Pangenome variation graphs model the mutual alignment of collections of DNA sequences.
A set of pairwise alignments implies a variation graph, but there are no scalable methods to generate such a graph from these alignments.
Existing related approaches depend on a single reference, a specific ordering of genomes, or a \textit{de Bruijn} model based on a fixed $k$-mer length. %, and may not describe the base-level alignment of genomes into the pangenome.
A scalable, self-contained method to build pangenome graphs without such limitations would be a key step in pangenome construction and manipulation pipelines. \\
\textbf{Results:}
We design the \textit{seqwish} algorithm, which builds a variation graph from a set of sequences and alignments between them.
We first transform the alignment set into an implicit interval tree.
To build up the variation graph, we query this tree-based representation of the alignments to reduce transitive matches into single DNA segments in a sequence graph.
By recording the mapping from input sequence to output graph, we can trace the original paths through this graph, yielding a pangenome variation graph.
We present an implementation that operates in external memory, using disk-backed data structures and lock-free parallel methods to drive the core graph induction step.
We demonstrate that our method scales to very large graph induction problems by applying it to build pangenome graphs for several species. \\
\textbf{Availability:}
\textit{seqwish} is published as free software under the MIT open source license.
Source code and documentation are available at \url{https://github.com/ekg/seqwish}.
\textit{seqwish} can be installed via Bioconda \url{https://bioconda.github.io/recipes/seqwish/README.html} or GNU Guix \url{https://github.com/ekg/guix-genomics/blob/master/seqwish.scm}. \\
\textbf{Contact:} \href{egarris5@uthsc.edu}{egarris5@uthsc.edu} \\
%\textbf{Supplementary information:} Supplementary data are available at \textit{Bioinformatics} online.
}
